---
title: "Take Home Ex 4.1 - Project Charts"
author: "Zachary Wong"
date: 03.18.2024
date-modified: "last-modified"
toc: true
execute: 
  eval: true
  echo: true
  warning: false
editor: visual
code-overflow: wrap
website:
  back-to-top-navigation: true
---

# Installing and Launching R Packages

```{r}
pacman::p_load(shiny, tidyverse, shinydashboard,dplyr,
               spatstat, spdep,
               lubridate, leaflet,
               plotly, DT, viridis,
               ggplot2, sf, tmap, readr,
               scales, ggthemes, gridExtra,
               knitr, data.table,
               CGPfunctions, ggHoriPlot, 
               patchwork, ggiraph, vcd, vcdExtra,
               ggstatsplot, ggmosaic, FunnelPlotR,
               knitr)
```

# Data Wrangling

## Importing Dataset

```{r}
Myanmar <- read_csv("data/2010-01-01-2023-12-31-Southeast_Asia-Myanmar.csv")
```

## Adjusting Attributes

```{r}
Myanmar <- Myanmar %>%
  mutate(year =factor(year))

Myanmar$event_date <- dmy(Myanmar$event_date)

```

## Reducing Columns

```{r}
Myanmar_final <- Myanmar %>%
  select(-time_precision, -geo_precision, -source_scale, -timestamp, -tags)
```

## Geo-data Correction

```{r}
ACLED_MMR_1 <- Myanmar_final %>%
  mutate(admin1 = case_when(
    admin1 == "Bago-East" ~ "Bago (East)",
    admin1 == "Bago-West" ~ "Bago (West)",
    admin1 == "Shan-North" ~ "Shan (North)",
    admin1 == "Shan-South" ~ "Shan (South)",
    admin1 == "Shan-East" ~ "Shan (East)",
    TRUE ~ as.character(admin1)
  ))
```

```{r}
ACLED_MMR_1 <- Myanmar_final %>%
  mutate(admin2 = case_when(
    admin2 == "Yangon-East" ~ "Yangon (East)",
    admin2 == "Yangon-West" ~ "Yangon (West)",
    admin2 == "Yangon-North" ~ "Yangon (North)",
    admin2 == "Yangon-South" ~ "Yangon (South)",
    admin2 == "Mong Pawk (Wa SAD)" ~ "Tachileik",
    admin2 == "Nay Pyi Taw" ~ "Det Khi Na",
    admin2 == "Yangon" ~ "Yangon (West)",
    TRUE ~ as.character(admin2)
  ))
```

# EDA

## Summary of Incidents

```{r}
Summary_Data <- ACLED_MMR_1 %>%
  group_by(year,event_type) %>%
  summarize(
    Total_incidents = n(),
    Total_Fatalities = sum(fatalities, na.rm=TRUE)
  )
```

### Distribution of Incidents Across Years by *event_type*

```{r}
#| fig-width: 14
#| fig-height: 20

gg1 <- ggplot(Summary_Data,
       aes(x = year,
           y = Total_incidents,
           fill = event_type)) +
  geom_col_interactive(aes(tooltip = year,
                           data_id = year),
                       alpha = 0.5) +
  geom_text(aes(label = Total_incidents), 
            vjust = 0.5, 
            color = "black",
            size = 2,
            check_overlap = TRUE,
            position = "dodge") +
  facet_wrap(~event_type, 
             ncol =2) +
  theme_minimal() +
  labs(y = "Total Number of Incidents", x = "Years") +
  theme(
    panel.grid.major.y = element_line(color = "pink", linetype = 2),
    strip.background = element_rect(fill = "black"),
    axis.text.x = element_blank(),
    strip.text = element_text(colour = "white"),
    legend.position = "none")

girafe(                                  
  ggobj = gg1,                             
  width_svg = 6,                         
  height_svg = 6*0.618)
  
```

### Distribution of Fatalities Across Years by *event_type*

```{r}
#| fig-width: 14
#| fig-height: 20

gg2 <- ggplot(Summary_Data,
       aes(x = year,
           y = Total_Fatalities,
           fill = event_type)) +
  geom_col_interactive(aes(tooltip = year,
                           data_id = year),
                       alpha = 0.5) +
  geom_text(aes(label = Total_incidents), 
            vjust = 0.5, 
            color = "black",
            size = 2,
            check_overlap = TRUE,
            position = "dodge") +
  facet_wrap(~event_type,
             ncol =2) +
  theme_minimal() +
  labs(y = "Total Number of Fatalities", x = "Years") +
  theme(
    panel.grid.major.y = element_line(color = "pink", linetype = 2),
    strip.background = element_rect(fill = "black"),
    strip.text = element_text(colour = "white"),
    axis.text.x = element_blank(),
    legend.position = "none")

girafe(                                  
  ggobj = gg2,                             
  width_svg = 6,                         
  height_svg = 6*0.618)

```

### Distribution of Incidents by Region

```{r}
Region_Summary <- ACLED_MMR_1 %>%
  group_by(country, admin1, admin2, admin3, event_type, disorder_type) %>%
  summarize(
    Total_incidents = n(),
    Total_Fatalities = sum(fatalities, na.rm=TRUE)
  )
```

```{r}
#| fig-width: 14
#| fig-height: 20


gg3 <- ggplot(Region_Summary,
       aes(x = admin1,
           y = Total_incidents,
           fill = event_type)) +
  geom_col_interactive(aes(tooltip = admin1,
                           data_id = admin1),
                       alpha = 0.5) +
  geom_text(aes(label = Total_incidents), 
            vjust = 0.5, 
            color = "black",
            size = 2,
            check_overlap = TRUE,
            position = "dodge") +
  facet_wrap(~event_type,
             ncol =2) +
  theme_minimal() +
  labs(y = "Total Number of Incidents", x = "Regions") +
  theme(
    panel.grid.major.y = element_line(color = "pink", linetype = 2),
    axis.text.x = element_blank(),
    strip.background = element_rect(fill = "black"),
    strip.text = element_text(colour = "white"),
    legend.position = "none")

girafe(                                  
  ggobj = gg3,                             
  width_svg = 6,                         
  height_svg = 6*0.618)

```

### Distribution of Fatalities by Region

```{r}
#| fig-width: 14
#| fig-height: 20


gg4 <- ggplot(Region_Summary,
       aes(x = admin1,
           y = Total_Fatalities,
           fill = event_type)) +
  geom_col_interactive(aes(tooltip = admin1,
                           data_id = admin1),
                       alpha = 0.5) +
  geom_text(aes(label = Total_Fatalities), 
            vjust = 0.5, 
            color = "black",
            size = 2,
            check_overlap = TRUE,
            position = "dodge") +
  facet_wrap(~event_type,
             ncol =2) +
  theme_minimal() +
  labs(y = "Total Number of Fatalities", x = "Regions") +
  theme(
    panel.grid.major.y = element_line(color = "pink", linetype = 2),
    axis.text.x = element_blank(),
    strip.background = element_rect(fill = "black"),
    strip.text = element_text(colour = "white"),
    legend.position = "none")

girafe(                                  
  ggobj = gg4,                             
  width_svg = 6,                         
  height_svg = 6*0.618)
```

# CDA

## Summary of Event Types by Fatalities (Total Count)

```{r}
ggbetweenstats(ACLED_MMR_1,
               x= event_type,
               y= fatalities)
```

## Summary of Event Types by Fatalities (Total Count), Grouped by Region

```{r}
#| fig-width: 14
#| fig-height: 25


grouped_ggbetweenstats(ACLED_MMR_1, 
              x = event_type, 
              y = fatalities,
              grouping.var = admin1,
              type = "np",
              pairwise.display = "s",
              pairwise_comparisons = TRUE,
              output = "plot")
```

## Summary of Event Types Across Years, Incidents and Fatalities

### Incidents

```{r}
ggbetweenstats(Summary_Data,
               x= event_type,
               y= Total_incidents)
```

### Fatalities

```{r}
ggbetweenstats(Summary_Data,
               x= event_type,
               y= Total_Fatalities)
```

## Mosaic Plot (Incident Counts)

### By Year and Event Type

```{r}
#| fig-width: 12
#| fig-height: 12


vcd::mosaic(~year + event_type,  data = ACLED_MMR_1, gp = shading_max)
```

### By Region and Event Type

```{r}
#| fig-width: 12
#| fig-height: 12


vcd::mosaic(~ admin1 + event_type,  data = ACLED_MMR_1, gp = shading_max)
```

## Using geom_mosiac

### Fatalities by Event Type and Region

```{r}
#| fig-width: 16
#| fig-height: 12
gg5 <- ggplot(Region_Summary) +
  geom_mosaic(aes(weight = Total_Fatalities,
                  x = (product(event_type, country)), fill = admin1)) +
  labs(x = "Myanmar",
       fill = "Regions") +
  ## guides(fill = "none") + to remove legend if not required
  theme(
    axis.text.x = element_blank(),
    axis.title.y = element_blank(),
    axis.ticks.x = element_blank()
  ) 



ggplotly(gg5)
```

# FunnelPlotR

```{r}
Region_Summary_Test <- ACLED_MMR_1 %>%
  filter(year == 2023)

Region_Summary_Test <- Region_Summary_Test %>%
  group_by(country, admin1, admin2, admin3, event_type) %>%
  summarize(
    Total_incidents = n(),
    Total_fatalities = sum(fatalities)
  )

Region_Summary_Test <- Region_Summary_Test %>%
  mutate(rate = Total_fatalities / Total_incidents) %>%
  mutate(rate.se = sqrt((rate*(1-rate))/ (Total_incidents))) %>%
  filter(rate>0)
```

```{r}
fit.mean <- weighted.mean(Region_Summary_Test$rate, 1/Region_Summary_Test$rate.se^2)
```

```{r}
number.seq <- seq(1, max(Region_Summary_Test$Total_incidents), 1)
number.ll95 <- fit.mean - 1.96 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
number.ul95 <- fit.mean + 1.96 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
number.ll999 <- fit.mean - 3.29 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
number.ul999 <- fit.mean + 3.29 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
dfCI <- data.frame(number.ll95, number.ul95, number.ll999, 
                   number.ul999, number.seq, fit.mean)
```

```{r}
p <- ggplot(Region_Summary_Test, aes(x = Total_incidents, y = rate)) +
  geom_point(alpha=0.4) +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ll95), 
            size = 0.4, 
            colour = "grey40", 
            linetype = "dashed") +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ul95), 
            size = 0.4, 
            colour = "grey40", 
            linetype = "dashed") +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ll999), 
            size = 0.4, 
            colour = "grey40") +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ul999), 
            size = 0.4, 
            colour = "grey40") +
  geom_hline(data = dfCI, 
             aes(yintercept = fit.mean), 
             size = 0.4, 
             colour = "grey40")

p
```

```{r}
funnel_plot(
  numerator = Region_Summary_Test$Total_incidents,
  denominator = Region_Summary_Test$Total_fatalities,
  group = Region_Summary_Test$admin2,
  data_type = "PR", #<<
  xrange = c(0,1500), #<<
  yrange = c(0,20) #<<


)
```
